---
title: "Activity monitoring data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## download and unzip data

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "./repdata%2Fdata%2Factivity.zip")
unzip("./repdata%2Fdata%2Factivity.zip")
activity <- read.csv("./activity.csv")
```

## Total number of steps taken each day

#### Calculate the total number of steps taken per day and make a histogram.
```{r}
step_d_t <- with(activity, tapply(steps, date, sum, na.rm=TRUE))
hist(step_d_t,xlab="steps",main="Total number of steps taken each day")

```
 
#### Calculate and the mean and median of the total number of steps taken per day.
```{r}
mean <- mean(step_d_t)
median <- median(step_d_t)
```
Mean of the total number of steps taken per day is `r mean`.
Median of the total number of steps taken per day is `r median`.

## The average daily activity pattern

#### Make a time series plot of the average number of steps taken across the 5-minute intervals.
```{r}
step_i_m <- with(activity, tapply(steps, interval, mean, na.rm=TRUE))
plot(names(step_i_m),step_i_m, type= "l",xlab = "interval", ylab = "average steps")
```

#### Calculate the 5-minute interval on average across contains the maximum number of steps.
```{r}
max <- names(which.max(step_i_m))
```
The 5-minute interval that on average contains the maximum number of steps is `r max`.

## Imputing missing values

#### Calculate the total number of missing values in the dataset.
```{r}
na <- sum(!complete.cases(activity))
```
There are `r na` missing values in the dataset.
 
#### Fill in all of the missing values in the dataset with the mean for that 5-minute interval.
```{r}
activity$step_m <- step_i_m
for (i in 1:17568){
    if (is.na(activity[i,1])==TRUE){
        activity[i,1] <- activity[i,4]
    }
}
activity <- activity[,1:3]
head(activity)
```

#### Make a histogram of the total number of steps taken each day after missing values are imputed.
```{r}
step_d_t2 <- with(activity, tapply(steps, date, sum))
hist(step_d_t2,xlab="steps",main="Total number of steps taken each day")
```

#### Calculate the mean and median total number of steps taken per day after missing values are imputed.
```{r}
mean2 <- mean(step_d_t2)
median2 <- median(step_d_t2)
```
Mean of the total number of steps taken per day is `r mean2`.
Median of the total number of steps taken per day is `r median2`.

## Panel plot of weekdays and weekends

#### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
Sys.setlocale("LC_ALL","English")
activity$date <- as.Date(activity$date, format="%Y-%m-%d")
activity$weekdays <- as.factor(weekdays(activity$date))
library("dplyr")
activity <- activity %>% mutate(w = ifelse(weekdays == "Saturday"|weekdays == "Sunday", "weekend", "weekday"))
head(activity)
```

#### Make a panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.
```{r}
activity1 <- filter(activity, w == "weekday")
activity2 <- filter(activity,w == "weekend")
step_i_m1 <- with(activity1, tapply(steps, interval, mean))
step_i_m2 <- with(activity2, tapply(steps, interval, mean)) 
par(mfrow = c(2,1),mar = c(4, 5, 3, 5))
plot(names(step_i_m1),step_i_m1, type= "l",xlab = "interval", ylab = "average steps", main = "weekdays")
plot(names(step_i_m2),step_i_m2, type= "l",xlab = "interval", ylab = "average steps", main = "weekends")
```